An Analysis of Performance Bottlenecks in MRI Pre-Processing

### 2024-05-22 paper/paper.tex
Small edits in conclusion

\section{Conclusion}
In this paper, we profiled several MRI pre-processing pipelines used in the popular fMRIPrep tool: ANTs brainExtraction, ANTs registration, FSL FAST, FSL FLIRT, FSL MCFLIRT, FreeSurfer recon-all. We observed that few functions contribute to a large amount of CPU time which offers opportunities to optimize a practical amount of functions to potentially obtain significant speed-up values. We found that most pipelines suffered from memory bound. Linear interpolation was the primary bottleneck for these pipelines. We discovered a bug in ITK which leads the implementation of ANTs registration in single-precision to use double precision in practice. We also discovered a potential bug in FreeSurfer recon-all which limits the benefit from multi-threading with OpenMP. Lastly, we discussed the challenge of profiling long-running pipelines with VTune, due to computational resource requirements.

@@

Future optimization efforts could focus on reduced precision techniques, to reduce memory bound and accelerate interpolation computation. We note that the use of reduced precision is not trivial though, as seen with the ANTs bug. Therefore, careful attention would be required to find a balance between performance and accuracy. We also suggest that future work study the different configurations of OpenMP scheduling in FreeSurfer recon-all, to improve parallel efficiency. This would lead to faster executions time and higher CPU usage in HPC allocations.

>>>>

\section{Conclusion}
In this paper, we profiled several MRI pre-processing pipelines used in the popular fMRIPrep tool: ANTs brainExtraction, ANTs registration, FSL FAST, FSL FLIRT, FSL MCFLIRT, FreeSurfer recon-all. We observed that few functions contribute to a large amount of CPU time which offers opportunities to optimize a practical amount of functions to potentially obtain significant speed-up values. We found that most pipelines suffered from memory bound, and that linear interpolation was the primary time bottleneck. We discovered a bug in ITK which leads the implementation of ANTs registration in single-precision to use double precision instead. We also discovered a potential bug in FreeSurfer recon-all which limits the benefit from multi-threading with OpenMP. Lastly, we discussed the challenge of profiling long-running pipelines with VTune, due to computational resource requirements.

@@

Future optimization efforts could focus on reduced precision techniques, to reduce memory bound and accelerate interpolation computation. We note that the use of reduced precision is not trivial, as seen with the ANTs bug. Therefore, careful attention would be required to find a balance between performance and accuracy. We also suggest that future work study the different configurations of OpenMP scheduling in FreeSurfer recon-all, to improve parallel efficiency. This would lead to faster executions time and higher CPU usage in HPC allocations.
